### 2015-05-22 reports/paper/experiment.tex
paper update

To test localization accuracy, an one meter by one meter grid was set up and the arrays are placed at the top left and top right corners of the grid. Fig~\ref{fig:setup_point} shows a picture of the setup. A total of $32$ positions are chosen uniformly in this region where microphone data is recorded.

\subsubsection{Movement tracking}
To test how well the arrays track movement, we mounted a rotating disk $40$ centimeter in diameter onto the grid at $(x=0$ cm$, y=-30$ cm$)$. Fig~\ref{fig:setup_circle} shows a picture of the sestup. A sound source is placed on the edge of rotating disk and the arrays localizes the sound source as it rotates in a circle. In this experiment, we tested how accuracy changes with:
\begin{itemize}
\item different window sizes
\item different audio sources
\item different movement tracking filters
\item different movement speeds
\end{itemize}

To test how different sound sources impact localization quality, we conducted three experiments on the same movement track with three different sound sources:

@@

\item[Music A] A music that has normal audio amplitude throughout experiment period was chosen. ``Honest Eyes'' by Black Tide was the music used.

\item[Music B] A music with intermittent low amplitude sections was chosen. ``Canon'' was the music used.

@@

To test how sound source movement speed affects localization quality, each of three experiments were conducted at two different speeds:
\begin{description}
\item[Normal] An angular speed of $0.5$ rad/s was maintained, which translates to linear speed of $10$ cm/s.
\item[Fast] An angular speed of $1.0$ rad/s was maintained, which translates to linear speed of $20$ cm/s.
\end{description}

For each experiment conducted, two different movement filters are tests:
\begin{description}[\IEEEsetlabelwidth{Very long label}\IEEEusemathlabelsep]
\item[Averaging filter] localization for past $0.5$ seconds are averaged and outputed as current estimate.
\item[Kalman filter] A 2nd order Kalman filtering is used.
\end{description}

\subsection{Results}
\subsubsection{Point localization}

To test how accuracy varies with window size, the algorithm is fed with recorded microphone data of different window size. Fig~\ref{fig:accuracy_vs_window} shows how accuracy changes with window length for three GCC algorithms. The error decreases as window size increases and plateaus after window size exceeds around $10$ millisecond. The lowest error achieved is $2.53$ centimeters. It is achieved when window size is set to 12 millisecond and GCC\_PHAT is used for TDOA estimation.

Although accuracy improves with window length, the calculation time also increases with window length. The part of calculation that depends on window length is using cross correlation to estimate TDOA. cross correlation can be calculated with FFT and the runtime is of order $O(N\log N)$. We measured how the computation time varies with window length and Figure~\ref{fig:speed_vs_window} shows the result. The runtime increases approximately linearly in the window size region of interest.

We also calculated the localization error for each tested point in the region. Figure~\ref{fig:error_distribution} shows a heatmap of the error distribution inside the grid. The error is below $3$ cm for most areas inside the region. There is one error spike in the mid-left region and we contribute this to audio source misplacecment becuase the error is fairly low and consistent around that spike region.

@@

\begin{figure}[]
\centering
\includegraphics[width=0.3\textwidth]{error_distribution.eps}
\caption{error distribution in the grid}
\label{fig:error_distribution}
\end{figure}

@@

Fig~\ref{fig:err_wn} gives an intuitive representation of how accuracy changes with window size. When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate TDOA. The localization is noisy. As window size increases, the localization converges to the shape of ground truth circle. Fig~\ref{fig:err_win_circle} shows how the error changes with window size. The general trend is similar to that in point localization case. The error decreases as window size increases and plateaus after window size exceeds around $10$ milliseconds.

>>>>

To test localization accuracy, an one meter by one meter grid was set up and the arrays are placed at the top left and top right corners of the grid. Fig~\ref{fig:setup_point} shows a picture of the setup. A total of $32$ testing locations are chosen uniformly in this.

\subsubsection{Movement tracking}
To test how well the arrays track movement, we mounted a rotating disk $40$ centimeter in diameter onto the grid at $(x=0$ cm$, y=-30$ cm$)$. Fig~\ref{fig:setup_circle} shows a picture of the sestup. A sound source is placed on the edge of the rotating disk and the arrays localizes the sound source as it rotates in a circle. In this experiment, we tested how accuracy changes with:
\begin{itemize}
\item different window sizes
\item different audio sources
\item different movement tracking filters
\item different movement speeds
\end{itemize}

To test how different sound sources impact localization quality, we conducted the experiments with three different sound sources:

@@

\item[Music A] A randomly picked music that has normal audio amplitude throughout experiment period. ``Honest Eyes'' by Black Tide was the music used.

\item[Music B] A randomly picked music with intermittent low amplitude sections. ``Canon'' was the music used.

@@

To test how the movement speed of sound source affects localization quality, each experiment was conducted at two different speeds:
\begin{description}
\item[Normal] An angular speed of $0.5$ rad/s was maintained, which translates to a linear speed of $10$ cm/s.
\item[Fast] An angular speed of $1.0$ rad/s was maintained, which translates to a linear speed of $20$ cm/s.
\end{description}

For each experiment conducted, two different movement filters were tests:
\begin{description}[\IEEEsetlabelwidth{Very long label}\IEEEusemathlabelsep]
\item[Averaging filter] localization for past $0.5$ seconds are averaged and outputed as current estimate.
\item[Kalman filter] A 2nd order Kalman filtering is used.
\end{description}

\subsection{Results}
\subsubsection{Point localization}
To test how accuracy varies with window size, the algorithm is fed with microphone data of different window size. Fig~\ref{fig:accuracy_vs_window} shows how accuracy changes with the window size for GCC, GCC\_PHAT and GCC\_PHAT\_SQRT. The error decreases as window size increases and plateaus after the window size exceeds around $10$ millisecond. The lowest error achieved is $2.53$ centimeters. It is achieved when window size is set to 12 millisecond and GCC\_PHAT is used for TDOA estimation.

Although accuracy improves with the window size, calculation time also increases with it. The part of calculation that depends on the window size is using cross correlation for TDOA estimation. Cross correlation can be calculated with Fast Fourier Transform (FFT) and the runtime is of order $O(N\log N)$. We measured how the computation time varies with window size and Figure~\ref{fig:speed_vs_window} shows the result. The runtime increases approximately linearly in the window size region of interest.

We also calculated the localization error for each tested point in the region. Figure~\ref{fig:error_distribution} shows the error distribution inside the grid. The error is below $3$ cm for most areas inside the region. There is one error spike in the mid-left region and we contribute this to audio source misplacecment becuase the error is fairly low and consistent around that spike.

@@

\begin{figure}[]
\centering
\includegraphics[width=0.3\textwidth]{error_distribution.eps}
\caption{error distribution in the grid. Arrays are placed at $(-0.5$ m$, 0$ m$)$ and $(0.5$ m$, 0$ m$)$}
\label{fig:error_distribution}
\end{figure}

@@

Fig~\ref{fig:wn} gives an intuitive representation of how accuracy changes with window size. When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate TDOA. The localization is noisy. As window size increases, the localization converges to the shape of ground truth circle. Fig~\ref{fig:err_win_circle} shows how the error changes with window size. The general trend is similar to that in point localization case. Error decreases as the window size increases and plateaus after it exceeds around $10$ milliseconds.

### 2015-05-22 reports/paper/simulation.tex
paper update

\section{Array Architecture}
As was mentioned in the previous section, points with the same TDOA to two fixed locations form a hyperbola on a 2D plane. However, in practical systems we can only measure TDOA up to a precision. Therefore we look at all points with difference of distance close to some target value within measurement error $\epsilon$. This $\epsilon$ represents accuracy on the measurement of difference of distances, and in practice it is related to sampling rate and estimation of difference of arrival time. In this section we evaluate the impact of difference of distance estimation on localization accuracy.

@@

To see how precision affects localization accuracy, we simulated two microphones placed at: $M_1:(x=-10\mbox{ cm},y=0\mbox{ cm})$ and $M_2:(x=10\mbox{ cm},y=0\mbox{ cm})$. A test sound source is emitted at point $P$ which is $50$ centimeters away from the origin $(0,0)$. Fig~\ref{fig:sim_2_5} shows the region $R$ where all points $\hat P$ satisfy:
\[
R:\{\hat P: |(\hat P M_1 - \hat P M_2) - a|< 1 \mbox{ cm}\}
\]
, where $a$ is the difference of distance between $P M_1$ and $P M_2$:
\[
P M_1 - P M_2 = a
\]
Intuitively, points in $R$ have difference of distance very similar to each other. Looking at fig~\ref{fig:sim_2_5}, we can still see that $R$ has the shape of a hyperbola, but with an uncertainty region around it. The thickness of the uncertainty region is not uniform around the hyperbola, the farther away the point is, the larger the uncertainty region becomes. This indicates for the the same delta distance movement it will generate smaller difference of distance change when the source is farther away from the array. The size of the uncertianty region is also angle dependent: points closer to the line connecting microphones have larger region compared to points close to the line bisecting microphones.

@@

With more than two microphones, each pair of microphones generates a hyperbolic region and localization becomes finding the intersection of hyperbolic regions. The smaller the intersection region, the better the localization accuracy. To see how accuracy changes with array placement and sound source location, three microphones are placed at the three vertices of an $20$ cm equilateral triangle. An audio source is placed at $20$ cm away from the center of the array. Fig~\ref{fig:sim_3_2} shows the intersection of regions for $5$ different placement of the sound source. It can be seen that accuracy is decreases when sound source becomes close to the line connecting any two microphones. This observation is consistent with two microphone case, since points close to lines connecting microphones have a larger uncertainty region.

To see how sound source distance affects localization accuracy, the same simulation is carried out with the sound source moved from $20$ cm to $80$ cm away from the center of the array. Results are presented in fig~\ref{fig:sim_3_8}. Comparing with fig~\ref{fig:sim_3_2}, accuracy decreases as the distance to the array increases. This is also consistent with our observation in $2$ microphone case where source farther away would result in larger uncertainty region.

@@

Intersection area is a measure of the localization accuracy. To evaluate an array's accuracy in a region, we can place sound source at predetermined grid points in the region and look at the intersection area for each tested point in the grid. The center location of intersection region can be used as localization estimate to calculate localization error. Results for a few different microphone array configurations are presented in fig~\ref{fig:sim_hm}.

Fig~\ref{fig:sim_hm_3} shows the accuracy when microphones are placed at three vertices of a $20$ cm equilateral triangle. The region inside the array has good accuracy. However, for regions along the line of any two microphones, the accuracy drops significantly. Average error across the region is $18.6$ cm.

To evaluate how adding one microphone(without increasing array size) improves accuracy,  another microphone is added to the array at $(0,0)$. Result is presented in fig~\ref{fig:sim_hm_3_p1}. Addition of the new microphone only slightly improved the accuracy around the array region. Average error dropped from $18.6$ cm to $17.1$ cm. Regions near lines of microphones still have significantly larger uncertainty region.

To evaluate array size's impact on accuracy, the size of original array from fig~\ref{fig:sim_hm_3} is increased by a factor of $2$. The result is presented in fig~\ref{fig:sim_hm_3_2x}. The overall uncertainty area decreased across the region. Average error improved to $10.04$ cm.

In fig~\ref{fig:sim_hm_3_line}, three microphones are placed $10$ cm apart from each other on x-axis. Error heatmap shows high uncertainty on the x axis, and the overall accuracy is not as good as that with three microphones placed in a triangle. The average error is $55.05$ cm.

@@

To avoid the need to accurately place four microphones at far distances(as required by fig~\ref{fig:sim_hm_4}), we explored configuration with two arrays. Two $3$ microphone array are placed $1$ meter apart and the result is presented in fig~\ref{fig:sim_hm_2_array}.  The result indicates that this configuration has good accuracy when source is close to the arrays. Accuracy decreases as sound source moves outside of the one meter by one meter region. The average error is $2.60$ cm.

>>>>

\section{Array Architecture}
As was mentioned in the previous section, points with the same TDOA to two fixed locations form a hyperbola on a 2D plane. However, in practical systems we can only measure TDOA up to a precision. Therefore we look at all points with difference of distance close to some target value within measurement error $\epsilon$. This $\epsilon$ represents accuracy on the measurement of difference of distances, and in practice it is related to sampling rate and estimation techniques. In this section we evaluate the impact of difference of distance estimation on localization accuracy.

@@

To see how precision affects localization accuracy, we simulated two microphones placed at: $M_1:(x=-10\mbox{ cm},y=0\mbox{ cm})$ and $M_2:(x=10\mbox{ cm},y=0\mbox{ cm})$. A test sound source is emitted at point $P$ which is $50$ centimeters away from the origin $(0,0)$. Let $2a$ denote the difference of distance between $P$ and two microphones:
\[
2a = P M_1 - P M_2 
\]
, then fig~\ref{fig:sim_2_5} shows the region $R$ where all points have difference of distance to two microphones close to $a$ within $1$ cm:
\[
R=\{\hat P: |(\hat P M_1 - \hat P M_2) - 2a|< 1 \mbox{ cm}\}
\]
Intuitively, points in $R$ have difference of distance to two microphones very similar to each other. Looking at fig~\ref{fig:sim_2_5}, we can still see that $R$ has the shape of a hyperbola, but with an uncertainty region around it. The thickness of the uncertainty region is not uniform around the hyperbola, the farther away the point is, the larger the uncertainty region becomes. This indicates for the the same delta distance movement it will generate smaller difference of distance change when the source is farther away from the array. The size of the uncertianty region is also angle dependent: points closer to the line connecting microphones have larger region compared to points close to the line bisecting microphones.

@@

With more than two microphones, each pair of microphones generates a hyperbolic region and localization becomes finding the intersection of hyperbolic regions. The smaller the intersection region, the better the localization accuracy. To see how accuracy changes with array placement and sound source location, three microphones are placed at the three vertices of an $20$ cm equilateral triangle. An audio source is placed at $20$ cm away from the center of the array. Fig~\ref{fig:sim_3_2} shows the intersection of regions for $5$ different placement of the sound source. It can be seen that accuracy decreases when sound source becomes close to the line connecting any two microphones. This observation is consistent with the two microphone case, since points close to lines connecting microphones have a larger uncertainty region.

To see how sound source distance affects localization accuracy, the same simulation is carried out with the sound source moved from $20$ cm to $80$ cm away from the center of the array. Results are presented in fig~\ref{fig:sim_3_8}. Comparing with fig~\ref{fig:sim_3_2}, accuracy decreases as the distance to the array increases. This is also consistent with our observation in $2$ microphone case where sources farther away would result in larger uncertainty region.

@@

Each microphone pair generates a hyperbolic region, and the intersection area of these regions is a measure of the localization accuracy. To evaluate an array's accuracy in a region, we can place sound source at predetermined grid points in the region and look at the intersection area for each tested point in the grid. The center location of intersection region can be used as the localization estimate to calculate localization error. Results for a few different microphone array configurations are presented in fig~\ref{fig:sim_hm}.

Fig~\ref{fig:sim_hm_3} shows the accuracy when microphones are placed at the three vertices of a $20$ cm equilateral triangle. The region inside the array has good accuracy. However, for regions along the line connecting any two microphones, the accuracy drops significantly. Average error across the region is $18.6$ cm.

To evaluate how adding one microphone(without increasing the array size) improves accuracy, another microphone is added to the array at $(0,0)$. Result is presented in fig~\ref{fig:sim_hm_3_p1}. Addition of the new microphone only slightly improved the accuracy around the array region. Average error dropped from $18.6$ cm to $17.1$ cm. Regions near lines connecting microphones still have significantly larger uncertainty region.

To evaluate the array size's impact on accuracy, the size of the original array (as in fig~\ref{fig:sim_hm_3}) is increased by a factor of $2$. The result is presented in fig~\ref{fig:sim_hm_3_2x}. The overall uncertainty area decreased across the region. Average error improved to $10.04$ cm.

In fig~\ref{fig:sim_hm_3_line}, three microphones are placed $10$ cm apart from each other on the x-axis. Error heatmap shows high uncertainty on the x axis, and the overall accuracy is not as good as that with three microphones placed in a triangle. The average error is $55.05$ cm.

@@

To avoid the need to accurately place microphones at far distances(as required by fig~\ref{fig:sim_hm_4}), we explored configuration with two arrays. Two $3$ microphone array are placed $1$ meter apart and the result is presented in fig~\ref{fig:sim_hm_2_array}.  The result indicates that this configuration has good accuracy when source is close to the arrays. Accuracy decreases as sound source moves outside of the one meter by one meter region. The average error is $2.60$ cm.

### 2015-05-22 reports/paper/system.tex
paper update

\section{System Setup}
The end system has two arrays, each with three microphones mounted on vertices of a $20$ cm equilateral triangle. A micro-controller is also attached to one of the vertices. Fig~\ref{fig:setup_array} shows a picture of the array setup. Micro-controller used in this project is \emph{teensy 3.1}. It has $64$k RAM memory and the ADC is capable of sampling at $500$kHz. In this project, the micro-controller collects microphone data on all three channels for $12$ millisecond and then send the recorded data to a computer through USB port for localization.

@@

To handle uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation function~\ref{eq:gcc2} as a mapping from delay to liklihood. For each microphone array, we build a heatmap for the region. The intensity at each point represents the likelihood of it being the source. For each point on the grid, the theoretical TDOA to each microphone pair can be precomputed. Then the heatmap can be generated by going through points on the grid and perform a lookup using equation~\ref{eq:gcc2}. With three microphones $m_1,m_2,m_3$, there are three microphones pairs: $m_1m_2,m_1m_3,m_2m_3$. Theoretical TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$. Then the liklihood map $L(x,y)$ can be built as:
\begin{eqnarray*}
  L(x,y) &=& R_{m_1,m_2}(D_{m_1,m_2}(x,y)) + R_{m_1,m_3}(D_{m_1,m_3}(x,y)) \\
 & & +R_{m_2,m_3}(D_{m_2,m_3}(x,y)) 
\end{eqnarray*}
,where $R_{m_1,m_2}(\tau)$,$R_{m_1,m_3}(\tau)$, and $R_{m_2,m_3}(\tau)$ denote GCC output from microphone pairs $m_1m_2,m_1m_3,$ and $m_2m_3$.

Liklihood maps from two arrays can be combined into final liklihood map:
\begin{equation}\label{eq:combine_l}
L(x,y) = L_1(x,y) L_2(x,y)
\end{equation}
, where $L_1(x,y)$ and $L_2(x,y)$ represents the liklihood map from array $1$ and array $2$.

@@

To see the effect of accuracy improvement using multiple arrays, fig~\ref{fig:liklihood} shows a real life localization where the source is placed at $(0$ cm$,-30$ cm$)$ . It shows the individual liklihood map from each array and also the combined liklihood map according to equation~\ref{eq:combine_l}. Individual array gives accurate angle estimate, but has high uncertainty in source distance. It demonstrated that by combining estimates from two arrays the system is able to perform more accurate localization.

From a timing point of view, the micro-controller spends $12$ millisecond on sampling microphone data before sending the data to the computer for processing. Sending data through USB port takes another $18$ millisecond, and processing on computer takes around $50$ millisecond. Therefore, the totoal time lag between sound source and localization is around $80$ millisecond.

>>>>

\section{System Setup}
The end system has two arrays, each with three microphones mounted on the vertices of a $20$ cm equilateral triangle. A micro-controller is attached to one of the vertices. Fig~\ref{fig:setup_array} shows a picture of the array setup. The micro-controller used in this project is \emph{teensy 3.1}, which has $64$k RAM memory and the ADC is capable of sampling at $500$kHz~\cite{tdoa:micloc, sys:teensy}. In this project, the micro-controller collects microphone data on all three channels for $12$ millisecond and then sends the recorded data to a computer through the USB port for localization.

@@

To handle the uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation output~\ref{eq:gcc2} as a measure of the likelihood of different arrival time differences. For each microphone array, we build a heatmap for the region. The intensity at each point represents the likelihood of it being the source. For each point on the grid, the theoretical TDOA to each microphone pair can be precomputed. Then the heatmap can be generated by going through points on the grid and performing a lookup using equation~\ref{eq:gcc2}. With three microphones $m_1,m_2,$ and $m_3$, there are three microphone pairs: $m_1m_2,m_1m_3,$ and $m_2m_3$. The theoretical TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$. Then the liklihood map $L(x,y)$ can be built by superposing the liklihood from each microphone pair:
\begin{eqnarray*}
L(x,y) &=& R_{m_1,m_2}(D_{m_1,m_2}(x,y)) + R_{m_1,m_3}(D_{m_1,m_3}(x,y)) \\
 & & +R_{m_2,m_3}(D_{m_2,m_3}(x,y)) 
\end{eqnarray*}
,where $R_{m_1,m_2}(\tau)$,$R_{m_1,m_3}(\tau)$, and $R_{m_2,m_3}(\tau)$ denote GCC output from microphone pairs $m_1m_2,m_1m_3,$ and $m_2m_3$.

Liklihood maps from two arrays can be combined into the final liklihood map:
\begin{equation}\label{eq:combine_l}
L(x,y) = L_1(x,y) L_2(x,y)
\end{equation}
, where $L_1(x,y)$ and $L_2(x,y)$ represent the liklihood map from array $1$ and array $2$.

@@

To see the effect of accuracy improvement using multiple arrays, fig~\ref{fig:liklihood} shows a real life localization where the source is placed at $(0$ cm$,-30$ cm$)$. It shows the individual liklihood map from each array and also the combined liklihood map according to equation~\ref{eq:combine_l}. Individual arrays give accurate angle estimate, but have high uncertainty in distance estimate. The combined liklihood map demonstrated that by merging estimates from two arrays the system is able to perform more accurate localization.

From a timing point of view, the micro-controller spends $12$ millisecond on sampling the microphone data before sending it to a computer for processing. Sending the data through the USB port takes another $15$ millisecond, and processing on the computer takes around $50$ millisecond. Therefore, the totoal time lag between sound source and localization is around $80$ millisecond.
